### 2024-04-20 other/datasheet/datasheet_template.qmd
added the other folder

---
title: "Datasheet for 'A dataset'"
author: 
  - First author
  - Another author
thanks: "Code and data are available at: LINK."
date: "`r Sys.time()`"
date-format: "D MMMM YYYY"
abstract: "First sentence. Second sentence. Third sentence. Fourth sentence."
format: pdf
number-sections: true
bibliography: references.bib
---

Extract of the questions from @gebru2021datasheets.

**Motivation**

1. *For what purpose was the dataset created? Was there a specific task in mind? Was there a specific gap that needed to be filled? Please provide a description.*
    - The dataset was created to enable analysis of Australian politicians. We were unable to find a publicly available dataset in a structured format that had the biographical and political information on Australian politicians that was needed for modelling.
2. *Who created the dataset (for example, which team, research group) and on behalf of which entity (for example, company, institution, organization)?*
    - TBD
3. *Who funded the creation of the dataset? If there is an associated grant, please provide the name of the grantor and the grant name and number.*
    - TBD
4. *Any other comments?*
    - TBD

**Composition**

1. *What do the instances that comprise the dataset represent (for example, documents, photos, people, countries)? Are there multiple types of instances (for example, movies, users, and ratings; people and interactions between them; nodes and edges)? Please provide a description.*
	- TBD
2. *How many instances are there in total (of each type, if appropriate)?*
	- TBD
3. *Does the dataset contain all possible instances or is it a sample (not necessarily random) of instances from a larger set? If the dataset is a sample, then what is the larger set? Is the sample representative of the larger set (for example, geographic coverage)? If so, please describe how this representativeness was validated/verified. If it is not representative of the larger set, please describe why not (for example, to cover a more diverse range of instances, because instances were withheld or unavailable).*
	- TBD
4. *What data does each instance consist of? "Raw" data (for example, unprocessed text or images) or features? In either case, please provide a description.*
	- TBD
5. *Is there a label or target associated with each instance? If so, please provide a description.*
	- TBD
6. *Is any information missing from individual instances? If so, please provide a description, explaining why this information is missing (for example, because it was unavailable). This does not include intentionally removed information, but might include, for example, redacted text.*
	- TBD
7. *Are relationships between individual instances made explicit (for example, users' movie ratings, social network links)? If so, please describe how these relationships are made explicit.*
	- TBD
8. *Are there recommended data splits (for example, training, development/validation, testing)? If so, please provide a description of these splits, explaining the rationale behind them.*
	- TBD
9. *Are there any errors, sources of noise, or redundancies in the dataset? If so, please provide a description.*
	- TBD
10. *Is the dataset self-contained, or does it link to or otherwise rely on external resources (for example, websites, tweets, other datasets)? If it links to or relies on external resources, a) are there guarantees that they will exist, and remain constant, over time; b) are there official archival versions of the complete dataset (that is, including the external resources as they existed at the time the dataset was created); c) are there any restrictions (for example, licenses, fees) associated with any of the external resources that might apply to a dataset consumer? Please provide descriptions of all external resources and any restrictions associated with them, as well as links or other access points, as appropriate.*
	- TBD
11. *Does the dataset contain data that might be considered confidential (for example, data that is protected by legal privilege or by doctor-patient confidentiality, data that includes the content of individuals' non-public communications)? If so, please provide a description.*
	- TBD
12. *Does the dataset contain data that, if viewed directly, might be offensive, insulting, threatening, or might otherwise cause anxiety? If so, please describe why.*
	- TBD
13. *Does the dataset identify any sub-populations (for example, by age, gender)? If so, please describe how these subpopulations are identified and provide a description of their respective distributions within the dataset.*
	- TBD
14. *Is it possible to identify individuals (that is, one or more natural persons), either directly or indirectly (that is, in combination with other data) from the dataset? If so, please describe how.*
	- TBD
15. *Does the dataset contain data that might be considered sensitive in any way (for example, data that reveals race or ethnic origins, sexual orientations, religious beliefs, political opinions or union memberships, or locations; financial or health data; biometric or genetic data; forms of government identification, such as social security numbers; criminal history)? If so, please provide a description.*
	- TBD
16. *Any other comments?*
	- TBD

**Collection process**

1. *How was the data associated with each instance acquired? Was the data directly observable (for example, raw text, movie ratings), reported by subjects (for example, survey responses), or indirectly inferred/derived from other data (for example, part-of-speech tags, model-based guesses for age or language)? If the data was reported by subjects or indirectly inferred/derived from other data, was the data validated/verified? If so, please describe how.*
	- TBD
2. *What mechanisms or procedures were used to collect the data (for example, hardware apparatuses or sensors, manual human curation, software programs, software APIs)? How were these mechanisms or procedures validated?*
	- TBD
3. *If the dataset is a sample from a larger set, what was the sampling strategy (for example, deterministic, probabilistic with specific sampling probabilities)?*
	- TBD
4. *Who was involved in the data collection process (for example, students, crowdworkers, contractors) and how were they compensated (for example, how much were crowdworkers paid)?*
	- TBD
5. *Over what timeframe was the data collected? Does this timeframe match the creation timeframe of the data associated with the instances (for example, recent crawl of old news articles)? If not, please describe the timeframe in which the data associated with the instances was created.*
	- TBD
6. *Were any ethical review processes conducted (for example, by an institutional review board)? If so, please provide a description of these review processes, including the outcomes, as well as a link or other access point to any supporting documentation.*
	- TBD
7. *Did you collect the data from the individuals in question directly, or obtain it via third parties or other sources (for example, websites)?*
	- TBD
8. *Were the individuals in question notified about the data collection? If so, please describe (or show with screenshots or other information) how notice was provided, and provide a link or other access point to, or otherwise reproduce, the exact language of the notification itself.*
	- TBD
9. *Did the individuals in question consent to the collection and use of their data? If so, please describe (or show with screenshots or other information) how consent was requested and provided, and provide a link or other access point to, or otherwise reproduce, the exact language to which the individuals consented.*
	- TBD
10. *If consent was obtained, were the consenting individuals provided with a mechanism to revoke their consent in the future or for certain uses? If so, please provide a description, as well as a link or other access point to the mechanism (if appropriate).*
	- TBD
11. *Has an analysis of the potential impact of the dataset and its use on data subjects (for example, a data protection impact analysis) been conducted? If so, please provide a description of this analysis, including the outcomes, as well as a link or other access point to any supporting documentation.*
	- TBD
12. *Any other comments?*
	- TBD

**Preprocessing/cleaning/labeling**

1. *Was any preprocessing/cleaning/labeling of the data done (for example, discretization or bucketing, tokenization, part-of-speech tagging, SIFT feature extraction, removal of instances, processing of missing values)? If so, please provide a description. If not, you may skip the remaining questions in this section.*
	- TBD
2. *Was the "raw" data saved in addition to the preprocessed/cleaned/labeled data (for example, to support unanticipated future uses)? If so, please provide a link or other access point to the "raw" data.*
	- TBD
3. *Is the software that was used to preprocess/clean/label the data available? If so, please provide a link or other access point.*
	- TBD
4. *Any other comments?*
	- TBD

**Uses**

1. *Has the dataset been used for any tasks already? If so, please provide a description.*
	- TBD
2. *Is there a repository that links to any or all papers or systems that use the dataset? If so, please provide a link or other access point.*
	- TBD
3. *What (other) tasks could the dataset be used for?*
	- TBD
4. *Is there anything about the composition of the dataset or the way it was collected and preprocessed/cleaned/labeled that might impact future uses? For example, is there anything that a dataset consumer might need to know to avoid uses that could result in unfair treatment of individuals or groups (for example, stereotyping, quality of service issues) or other risks or harms (for example, legal risks, financial harms)? If so, please provide a description. Is there anything a dataset consumer could do to mitigate these risks or harms?*
	- TBD
5. *Are there tasks for which the dataset should not be used? If so, please provide a description.*
	- TBD
6. *Any other comments?*
	- TBD

**Distribution**

1. *Will the dataset be distributed to third parties outside of the entity (for example, company, institution, organization) on behalf of which the dataset was created? If so, please provide a description.*
	- TBD
2. *How will the dataset be distributed (for example, tarball on website, API, GitHub)? Does the dataset have a digital object identifier (DOI)?*
	- TBD
3. *When will the dataset be distributed?*
	- TBD
4. *Will the dataset be distributed under a copyright or other intellectual property (IP) license, and/or under applicable terms of use (ToU)? If so, please describe this license and/ or ToU, and provide a link or other access point to, or otherwise reproduce, any relevant licensing terms or ToU, as well as any fees associated with these restrictions.*
	- TBD
5. *Have any third parties imposed IP-based or other restrictions on the data associated with the instances? If so, please describe these restrictions, and provide a link or other access point to, or otherwise reproduce, any relevant licensing terms, as well as any fees associated with these restrictions.*
	- TBD
6. *Do any export controls or other regulatory restrictions apply to the dataset or to individual instances? If so, please describe these restrictions, and provide a link or other access point to, or otherwise reproduce, any supporting documentation.*
	- TBD
7. *Any other comments?*
	- TBD

**Maintenance**

1. *Who will be supporting/hosting/maintaining the dataset?*
	- TBD
2. *How can the owner/curator/manager of the dataset be contacted (for example, email address)?*
	- TBD
3. *Is there an erratum? If so, please provide a link or other access point.*
	- TBD
4. *Will the dataset be updated (for example, to correct labeling errors, add new instances, delete instances)? If so, please describe how often, by whom, and how updates will be communicated to dataset consumers (for example, mailing list, GitHub)?*
	- TBD
5. *If the dataset relates to people, are there applicable limits on the retention of the data associated with the instances (for example, were the individuals in question told that their data would be retained for a fixed period of time and then deleted)? If so, please describe these limits and explain how they will be enforced.*
	- TBD
6. *Will older versions of the dataset continue to be supported/hosted/maintained? If so, please describe how. If not, please describe how its obsolescence will be communicated to dataset consumers.*
	- TBD
7. *If others want to extend/augment/build on/contribute to the dataset, is there a mechanism for them to do so? If so, please provide a description. Will these contributions be validated/verified? If so, please describe how. If not, why not? Is there a process for communicating/distributing these contributions to dataset consumers? If so, please provide a description.*
	- TBD
8. *Any other comments?*
	- TBD

\newpage

# References

### 2024-04-20 scripts/04-model.R
updated model file

@@ -30,7 +30,7 @@ shapiro.test(residuals(model))
 bptest(model)
 
 # Check for multicollinearity
-cor(cleaned_data$consumer_debt, cleaned_data$interest_rates)
+cor(analysis_data$consumer_debt, analysis_data$interest_rates)
 
 #### Save model ####
 saveRDS(

### 2024-04-20 paper/paper.qmd
added the paper

---
title: "Impact of Bank of Canada's Monetary Policy on Consumer Debt Levels"
author: 
  - Abdul Ahad Qureshi
thanks: "Code and data are available at: https://github.com/abdulahadq29/debt."
date: today
date-format: long
abstract: "This analysis investigates the relationship between changes in the Bank of Canada's interest rate policy and consumer debt levels in Canada. Using a dataset of quarterly financial flows and stocks of household credit market debt, the impact of short-term interest rates on credit market debt to disposable income is examined. A linear regression model showed an inverse relationship between short-term interest rates and consumer debt levels. This analysis helps us in understanding the role of fiscal tools in shaping consumer financial behavior."
format: pdf
number-sections: true
bibliograph: references.bib
---

```{r}
#| include: false
#| warning: false
#| message: false

library(ggplot2)
library(dplyr)
library(arrow)
library(here)
library(modelsummary)
#IMPORT ANALYSIS DATA
analysis_data <- read.csv(here("data", "analysis_data", "analysis_data.csv"))

```

# Introduction

In contemporary economies, central banks play a pivotal role in shaping economic conditions and maintaining financial stability through their monetary policy decisions. Among the various tools at their disposal, adjustments in interest rates stand out as one of the most influential mechanisms for influencing economic activity and financial markets. The Bank of Canada, as Canada's central bank, is entrusted with the responsibility of formulating and implementing monetary policy to achieve its dual mandate of price stability and sustainable economic growth. Central to the Bank of Canada's policy arsenal is the manipulation of short-term interest rates, primarily through its key policy rate, known as the overnight rate.

The relationship between central bank interest rate policy and various economic indicators has been the subject of extensive academic inquiry and policy debate. One area that has garnered significant attention is the impact of interest rate policy changes on consumer behavior, particularly in the realm of borrowing and household debt. Understanding how changes in interest rates affect consumer debt levels is crucial for policymakers, financial institutions, and individuals alike, as it has profound implications for financial stability, economic growth, and the welfare of households.

Against this backdrop, this paper seeks to investigate the relationship between changes in the Bank of Canada's interest rate policy and consumer debt levels in Canada. The estimand of our linear regression model quantifies the average effect that a one-unit change in the Bank of Canada’s short-term interest rates has on the consumer credit market debt to disposable income ratio in Canada.

The significance of this research question stems from several key factors. Firstly, consumer spending constitutes a significant portion of aggregate demand in the Canadian economy, making it a crucial driver of economic growth. Changes in consumer borrowing behavior, influenced by fluctuations in interest rates, can have ripple effects throughout the economy, affecting sectors such as housing, retail, and durable goods consumption. Secondly, household debt levels in Canada have been on the rise in recent years, reaching record highs relative to income levels. The sustainability of this debt burden and its implications for financial stability have become subjects of considerable public and policy concern. Thirdly, against the backdrop of economic uncertainty and global financial market volatility, the efficacy of monetary policy tools, including interest rate adjustments, in managing consumer behavior and mitigating risks associated with excessive debt accumulation warrants thorough examination.

To address these issues, this paper adopts a quantitative approach, leveraging empirical data on household debt levels and short-term interest rates in Canada. By analyzing the relationship between these variables over a specified time period, we aim to uncover insights into how changes in interest rates influence consumer borrowing decisions and overall debt dynamics. The findings of this study are expected to contribute to the existing body of literature on monetary policy transmission mechanisms and provide valuable insights for policymakers, financial institutions, and households navigating the complex landscape of debt and financial decision-making.

# Data {#sec-data}

## Data Management

Data analysis for this paper is performed in R [@citeR], and additional help is provided by libraries such as `dplyr` [@dplyr], `ggplot2` [@ggplot], `ggrepel` [@ggrepel], `tidyverse` [@thereferencecanbewhatever], `kableExtra` [@kableextra], `knitr` [@knitr], `haven` [@haven], 'readr' [@readr], `stargazer`[@stargazer], `psych` [@psych], `scales` [@scales], `RColorBrewer` [@rcolorbrewer], `gridExtra` [@gridextra], `here` [@here].

## Source

The empirical analysis conducted in this study relies on comprehensive datasets capturing key variables related to consumer debt levels and short-term interest rates in Canada. These datasets provide the empirical foundation for investigating the relationship between changes in the Bank of Canada's interest rate policy and consumer borrowing behavior.

The primary source of data on consumer debt levels is sourced from Statistics Canada. This dataset includes information on various metrics of household indebtedness, including aggregate measures of credit market debt to disposable income, consumer credit, mortgage liabilities, non-mortgage loans, and other relevant indicators. The data cover a substantial time span, allowing for a longitudinal analysis of trends and patterns in consumer borrowing behavior over time.

Complementing the dataset on consumer debt levels is information on short-term interest rates, obtained from FRED. This dataset comprises time-series data on the Bank of Canada's key policy rate, commonly referred to as the overnight rate or the target for the overnight rate. Additionally, the dataset may include data on other short-term interest rates, such as the prime lending rate set by commercial banks, providing a comprehensive view of interest rate movements in the Canadian financial system.

## Data Cleaning

The process involved removing outliers, handling missing values, and standardizing variable formats to facilitate robust analysis. The dates were converted to the year-month format. The two datasets were merged to create an unified dataset for analysis. The dataset was filtered for the "Credit market debt to disposable income" category. Finally, I renamed the last column in the original dataset from "IRSTCB01CAM156N" to "interest_rates" for the data to look better and be self-explanatory. Furthermore, the datasets were organized in a format conducive to statistical analysis, allowing for the application of econometric techniques to model the relationship between interest rate policy changes and consumer debt levels.

```{r}
#| label: fig-debt
#| fig-cap: "Trends in Credit Market Debt to Disposable Income"
#| echo: false
#| warning: false

# Convert REF_DATE to Date format if it isn't already
analysis_data$REF_DATE <- as.Date(analysis_data$REF_DATE)

# Aggregate data by REF_DATE, calculating the mean value for each date
analysis_data_agg <- analysis_data %>%
  group_by(REF_DATE) %>%
  summarise(consumer_debt = mean(consumer_debt, na.rm = TRUE)) %>%
  ungroup()  # Remove grouping for ggplot

# Plot the aggregated data
ggplot(analysis_data_agg, aes(x = REF_DATE, y = consumer_debt)) +
  geom_line() +  # Draw lines between points
  labs(title = "Trends in Credit Market Debt to Disposable Income",
       x = "Date",
       y = "Debt to Disposable Income Ratio") +
  theme_minimal()  
```

The plot visualizes the trends in the Credit Market Debt to Disposable Income ratio over time, spanning from 1990 to the present day. The x-axis represents the date, while the y-axis depicts the Debt to Disposable Income Ratio, a key metric used to assess the financial health of households.

From the early 1990s to around 2020, the curve exhibits a consistent upward trend, indicating a gradual increase in the Debt to Disposable Income Ratio over the years. This prolonged period of growth suggests a pattern of rising indebtedness among Canadian households relative to their disposable income levels. Factors contributing to this trend may include easier access to credit, low borrowing costs, demographic shifts, and changes in consumer spending patterns.

However, a notable inflection point occurs around the year 2020, marked by a sudden dip in the Debt to Disposable Income Ratio. This deviation from the previous upward trajectory suggests a significant shift in consumer borrowing behavior or macroeconomic conditions.

```{r}
#| label: fig-interest
#| fig-cap: "Trends in Short-term Interest Rates"
#| echo: false
#| warning: false

ggplot(analysis_data, aes(x = REF_DATE, y = interest_rates)) +
  geom_line(color = "blue") +
  labs(title = "Trends in Short-term Interest Rates",
       x = "Date",
       y = "Short-term Interest Rates") +
  theme_minimal()

```

The plot illustrates the trends in short-term interest rates over time, spanning from 1990 to the present day. The x-axis represents the date, while the y-axis depicts the level of short-term interest rates.

In the early 1990s, the plot indicates a period of relatively high short-term interest rates, suggesting a climate of tight monetary policy aimed at curbing inflationary pressures or addressing macroeconomic imbalances. However, from the mid-1990s onwards, there is a discernible downward trend in short-term interest rates, reflecting efforts by monetary authorities to stimulate economic growth or support financial stability through accommodative monetary policy measures.

Throughout the subsequent years, the plot shows smaller fluctuations in short-term interest rates, characterized by periodic ups and downs. These fluctuations may be attributed to various factors, including changes in economic conditions, shifts in inflation expectations, or adjustments in central bank policy settings in response to evolving economic circumstances.

# Model {#sec-model}

The goal of our modeling strategy is to quantify the relationship between changes in short-term interest rates and consumer debt levels. We employ a linear regression model to assess how variations in interest rates impact consumer borrowing behavior.

The linear regression model employed in this study aims to examine the relationship between changes in the Bank of Canada's interest rate policy and consumer debt levels in Canada. The model seeks to quantify the impact of fluctuations in short-term interest rates on the Credit Market Debt to Disposable Income ratio, a key indicator of household indebtedness.

The model's primary explanatory variable is the short-term interest rate, represented by the IRSTCB01CAM156N variable, sourced from the dataset on short-term interest rates. This variable serves as a proxy for changes in the Bank of Canada's monetary policy stance, capturing variations in borrowing costs and the availability of credit in the economy.

The response variable in the model is the Credit Market Debt to Disposable Income ratio, derived from the dataset on household debt levels. This ratio provides insights into the extent to which Canadian households are leveraging their disposable income to finance consumption and investment activities, reflecting their overall financial health and resilience to economic shocks.

The linear regression model assumes a linear relationship between the explanatory variable (short-term interest rates) and the response variable (Debt to Disposable Income ratio). By estimating the coefficients of the model, we seek to quantify the magnitude and direction of the relationship between changes in interest rates and consumer debt levels.

## Model set-up

```{=tex}
\begin{equation}
Y_{i} = \beta_{0} + \beta_{1}X_{i} + \epsilon_{i}
\end{equation}
```
In the Model:

-   $Y_{i}$ is the dependent variable representing the Credit Market Debt to Disposable Income ratio for each observation $i$.
- $X_{i}$is the independent variable representing the short-term interest rates for each observation $i$.
-   $\beta_{0}$ is the intercept of the regression line, indicating the expected level of $Y_{i}$ when $X_{i}$ is zero.
-   $\beta_{1}$ estimates the effect of the short-term interest rates on the Credit Market Debt to Disposable Income ratio.
- $\epsilon_{i}$ is the error term for each observation i,capturing the deviation of the observed $Y_{i}$ from the value predicted by the model.

## Model Justification

The choice of a linear regression model for analyzing the impact of the Bank of Canada’s interest rate policy on consumer debt levels is grounded in several key considerations. Firstly, linear regression is a widely-recognized statistical tool for assessing the strength and character of the relationship between two continuous variables. Given the quantitative nature of both interest rates and consumer debt levels, a linear model provides a straightforward, yet powerful, means to decipher the underlying patterns in the data. Its simplicity enables clear interpretation of results, such as understanding how a unit change in interest rates correlates with the magnitude of change in consumer debt levels. Secondly, the model’s simplicity should not be mistaken for a lack of rigor or complexity. Linear regression models are capable of capturing and illustrating complex economic phenomena when adequately specified and supported by robust data. By including a properly defined interest rate variable as the predictor, the model leverages data variability to illuminate the effect of monetary policy on household indebtedness. Thirdly, this approach aligns with economic theory, which postulates a linear or near-linear response of debt levels to changes in interest rates, at least within certain bounds. The assumption is that households adjust their borrowing in response to the cost of borrowing; as interest rates decrease, it becomes cheaper to borrow, encouraging higher levels of debt, and vice versa. This behavior is expected to follow a linear trend in the short to medium term, which is the focus of our analysis. Moreover, the linear regression model offers an entry point for further exploration. While it lays the foundation for understanding the direct relationship between interest rates and debt levels, it also allows for the addition of interaction terms and polynomial expansions in more complex iterations. This could be particularly useful in future research endeavors to explore potential non-linear dynamics or threshold effects as debt levels reach certain levels relative to income. Finally, the model’s robustness can be validated through various diagnostic tests, such as checks for homoscedasticity, normality of residuals, and multicollinearity, ensuring that the assumptions necessary for linear regression are met. The transparency with which a linear regression model can be subjected to such tests adds to the credibility and reliability of the findings. In conclusion, the linear regression model is an appropriate choice for our analysis, providing a methodologically sound, theoretically justified, and empirically robust framework for understanding how monetary policy impacts household debt levels in Canada.

# Results {#sec-results}

The results obtained from our linear regression model are summarized in @tbl-regression.

```{r}
#| label: tbl-regression
#| tbl-cap: "Linear Regression Results"
#| echo: false
#| eval: true
#| warning: false
#| message: false

# Fit linear regression model
debt_interest_model <- lm(consumer_debt ~ interest_rates, data = analysis_data)

modelsummary(debt_interest_model)
```

The significant coefficients obtained from the linear regression model provide valuable insights into the relationship between short-term interest rates and consumer debt levels in Canada. Let's delve deeper into the interpretation of these coefficients:

1.  **Intercept Coefficient (173.2100):**
    -   The intercept coefficient represents the expected value of the response variable (consumer debt levels) when the predictor variable (short-term interest rates) is zero.
    -   In this context, the intercept coefficient of 173.2100 suggests that when short-term interest rates are at zero, the estimated average consumer debt level is approximately 173.21 units. However, it's important to note that short-term interest rates are unlikely to be exactly zero in practice, so this interpretation is theoretical.
    -   The standard error associated with the intercept coefficient (2.9309) indicates the variability or uncertainty in estimating the intercept value.
2.  **Coefficient for Short-term Interest Rates (-8.9188):**
    -   The coefficient for short-term interest rates quantifies the estimated change in consumer debt levels for each unit increase in short-term interest rates.
    -   In this case, the coefficient of -8.9188 suggests that for every one-unit increase in short-term interest rates, the average consumer debt level is estimated to decrease by approximately 8.92 units.
    -   The negative sign of the coefficient indicates an inverse relationship between short-term interest rates and consumer debt levels. As interest rates increase, consumer debt levels tend to decrease, and conversely, as interest rates decrease, consumer debt levels tend to increase.
    -   The standard error associated with the coefficient (0.6482) reflects the uncertainty in estimating the precise magnitude of the relationship between interest rates and consumer debt levels.
3.  **Statistical Significance:**
    -   Both coefficients are highly statistically significant, as indicated by the p-values (p \< 0.001). This implies that the observed relationships between short-term interest rates and consumer debt levels are unlikely to have occurred by random chance alone.
    -   The high significance levels strengthen the confidence in the estimated coefficients and the validity of the relationship between interest rates and consumer debt levels.

The linear regression model results reveal significant coefficients for both the intercept and the predictor variable representing short-term interest rates (**`interest_rates`**). The intercept coefficient is estimated to be 173.2100 with a standard error of 2.9309, indicating the expected value of the response variable (consumer debt levels) when the predictor variable is zero. The coefficient for short-term interest rates is estimated to be -8.9188 with a standard error of 0.6482. This coefficient represents the estimated change in consumer debt levels for each unit increase in short-term interest rates. The negative sign of the coefficient suggests that as interest rates increase, consumer debt levels tend to decrease, and vice versa. Both coefficients are highly statistically significant (p \< 0.001), indicating a strong relationship between short-term interest rates and consumer debt levels.

The graph @fig-regression visualizes the correlation between consumer debt and short-term interest rates.

```{r}
#| label: fig-regression
#| tbl-cap: "Relationship between Consumer Debt Levels and Interest Rate"
#| echo: false
#| warning: false

# Load necessary libraries
library(tidyverse)
library(ggplot2)
library(readr) # if you are reading the data from a csv file

# Visualize the relationship between debt levels and interest rates
ggplot(analysis_data, aes(x = interest_rates, y = consumer_debt)) +
  geom_point() +
  geom_smooth(method = "lm", color = "blue") +
  labs(title = "Relationship between Debt Levels and Interest Rates",
       x = "Short-term Interest Rates",
       y = "Credit Market Debt to Disposable Income") +
  theme_minimal()

```

Our analysis indicates a significant inverse relationship between changes in short-term interest rates and consumer debt levels. As interest rates increase, consumer debt levels tend to decrease, and conversely, as interest rates decrease, consumer debt levels tend to increase. This finding underscores the importance of monetary policy decisions, particularly changes in interest rates, in influencing consumer borrowing behavior and overall debt accumulation.

# Discussion {#sec-discussion}

The implications of our findings extend well beyond the direct relationship between interest rate policy and consumer debt levels.

## For Policymakers ##

For economic policymakers, the significant inverse relationship uncovered in this study shows the powerful influence that the Bank of Canada's interest rate decisions have on household financial stability. In particular, our analysis reveals that consumer indebtedness is responsive to the cost of borrowing, which can be modulated via the overnight rate. This implies that interest rate adjustments could serve as a counterbalance to overheating credit markets, tempering excessive borrowing and potential bubbles, especially in real estate.his insight grants policymakers a powerful instrument to modulate economic activity and safeguard financial stability, yet it also imposes a considerable responsibility to wield this tool judiciously.

Interest rate decisions have traditionally been centered around controlling inflation and stabilizing prices, but our findings suggest these decisions also resonate deeply within the fabric of household economics. As such, policymakers must consider the broader impact of rate adjustments on consumer borrowing and spending behaviors. This is especially salient in the context of rising household debt, which poses a systemic risk to the economy. By carefully managing interest rates, the Bank of Canada can exert influence over the degree of leverage within the economy, mitigating the risks of both excessive debt accumulation during times of low rates and potential defaults during periods of rate hikes.

The data-driven approach of our analysis also advocates for a data-informed policymaking process. As consumer debt levels react to interest rate changes, the Bank of Canada and other institutions could harness such empirical evidence to refine their predictive models and policy frameworks. They can further employ advanced analytics to anticipate the potential ripple effects of their policy decisions across various sectors, from housing to consumer goods.

Moreover, this research highlights the need for policymakers to communicate their policy actions and rationale effectively to the public. Clear communication can help manage expectations and foster an environment where consumers are better informed about the implications of their financial decisions. As part of their mandate, policymakers could collaborate with educational institutions and non-profit organizations to promote financial literacy, ensuring that the nuances of monetary policy and its impact on everyday financial decisions are well understood by the broader population.

Furthemore, the impact of interest rates on consumer debt calls for a synchronized approach to policy formulation, where monetary, fiscal, and regulatory policies are harmoniously aligned. Coordination with fiscal authorities can amplify the intended effects of monetary adjustments, whether it is to stimulate economic growth or cool down overheating market segments. Regulatory bodies can ensure that the credit market remains robust and capable of withstanding interest rate fluctuations without precipitating undue stress on consumers.

## For Households ##

The findings from this study hold significant implications for households, shedding light on the tangible impact of central bank interest rate policies on their financial health. The demonstrated sensitivity of debt levels to changes in interest rates means that household financial strategies need to be adaptive and forward-looking.

For individual households, the rise and fall of interest rates directly influence their capacity to borrow and repay. Lower interest rates typically translate to cheaper loans, which can be an opportune time for households to secure mortgages or refinance existing debts. However, this apparent affordability can lead to increased borrowing, which, if unchecked, may result in heightened financial vulnerability. Households should be cautious not to over-leverage in enticing low-rate environments, keeping in mind that rates can increase, leading to higher repayment costs.

On the flip side, when interest rates rise, the cost of servicing debt increases. Households with variable-rate debts or those considering new loans must account for these higher costs. Those on fixed incomes or with tight budgets may find themselves particularly strained. Prudent financial management, such as prioritizing debt repayment and avoiding unnecessary borrowing during such periods, becomes even more crucial.

The relationship between debt levels and interest rates also has implications for long-term financial planning. Decisions around savings, investment, retirement planning, and even day-to-day budgeting must all factor in the potential for interest rate variability. Households must consider their exposure to interest rate risk and develop strategies to mitigate this. For instance, locking in a fixed-rate mortgage can shield from future rate hikes, while ensuring an emergency fund can provide a buffer against increased debt servicing costs.

Moreover, the study underscores the importance of financial literacy. A clear understanding of how monetary policy influences debt levels can empower consumers to make informed decisions. For example, awareness of the broader economic conditions and the Bank of Canada's policy stance can provide households with anticipatory guidance on whether to prioritize saving over spending or vice versa.

Households also stand to benefit from staying informed about economic forecasts and potential interest rate changes, which can influence major financial decisions such as purchasing property or taking on significant debt. By being proactive and staying informed, households can better position themselves to respond to the economic shifts driven by monetary policy.

# Limitations and Future Research {#sec-limitations}

While our study provides valuable insights into the relationship between interest rate policy and consumer debt levels, it is essential to acknowledge certain limitations. One limitation is the reliance on aggregate-level data, which may mask variations in borrowing behavior among different demographic groups and regions. Future research could explore these variations by analyzing micro-level data and examining how factors such as income, employment, and housing market conditions interact with interest rate policy to influence consumer debt dynamics. Additionally, further investigation into the mechanisms through which interest rate changes impact borrowing behavior, such as through mortgage rates, credit card rates, and other lending channels, could provide deeper insights into the transmission channels of monetary policy. Moreover, longitudinal studies tracking changes in consumer debt levels over time in response to interest rate policy changes could provide valuable information on the dynamics of household indebtedness and its implications for financial stability.

# References

### 2024-04-20 paper/paper.qmd
updated the paper

@@
 number-sections: true
-bibliograph: references.bib
+bibliography: references.bib
 ---
@@
 library(modelsummary)
+library(knitr)
 #IMPORT ANALYSIS DATA
@@
 
-Data analysis for this paper is performed in R [@citeR], and additional help is provided by libraries such as `dplyr` [@dplyr], `ggplot2` [@ggplot], `ggrepel` [@ggrepel], `tidyverse` [@thereferencecanbewhatever], `kableExtra` [@kableextra], `knitr` [@knitr], `haven` [@haven], 'readr' [@readr], `stargazer`[@stargazer], `psych` [@psych], `scales` [@scales], `RColorBrewer` [@rcolorbrewer], `gridExtra` [@gridextra], `here` [@here].
+Data analysis for this paper is performed in R [@citeR], and additional help is provided by libraries such as `dplyr` [@dplyr], `ggplot2` [@ggplot], `ggrepel` [@ggrepel], `tidyverse` [@thereferencecanbewhatever], `kableExtra` [@kableextra], `knitr` [@knitr], `haven` [@haven], `readr` [@readr], `RColorBrewer` [@rcolorbrewer], `gridExtra` [@gridextra], `here` [@here], `modelsummary` [@modelsummary]
 
@@
 These datasets provide the empirical foundation for investigating the relationship between changes in the Bank of Canada's interest rate policy and consumer borrowing behavior.
-
 The primary source of data on consumer debt levels is sourced from Statistics Canada.
-This dataset includes information on various metrics of household indebtedness, including aggregate measures of credit market debt to disposable income, consumer credit, mortgage liabilities, non-mortgage loans, and other relevant indicators.
-The data cover a substantial time span, allowing for a longitudinal analysis of trends and patterns in consumer borrowing behavior over time.
-
 Complementing the dataset on consumer debt levels is information on short-term interest rates, obtained from FRED.
-This dataset comprises time-series data on the Bank of Canada's key policy rate, commonly referred to as the overnight rate or the target for the overnight rate.
-Additionally, the dataset may include data on other short-term interest rates, such as the prime lending rate set by commercial banks, providing a comprehensive view of interest rate movements in the Canadian financial system.
 
@@
 Furthermore, the datasets were organized in a format conducive to statistical analysis, allowing for the application of econometric techniques to model the relationship between interest rate policy changes and consumer debt levels.
+
+## Data Measurement
+
+In our analysis, two primary metrics form the foundation of the dataset: consumer debt levels and interest rates.
+Consumer debt is captured as the total debt held by individuals within the domestic economy, encompassing various forms of debt such as credit card balances, student loans, auto loans, and mortgages.
+The data on consumer debt levels is measured quarterly and reported as an aggregate sum in millions of dollars.
+This comprehensive approach allows us to observe broader economic patterns and the impact of debt on financial stability.
+
+Interest rates are represented by the central bank’s benchmark short-term rate, which influences borrowing costs throughout the economy.
+These rates are recorded on a monthly basis and are presented as a percentage, reflecting the cost of borrowing.
+The interplay between consumer debt levels and interest rates is critical, as fluctuations in rates can significantly affect the ability of consumers to service their debts, thus impacting overall economic health.
+
+In this context, measurement accuracy is paramount.
+The data utilized in this study have been sourced from reputable financial databases and governmental reports, ensuring reliability and precision in the metrics employed.
+As economic indicators, both consumer debt and interest rates are subjected to rigorous methods of collection, involving checks and balances to validate their correctness.
+The data is then standardized to allow for an accurate comparative analysis over time, mitigating inconsistencies that may arise from raw financial figures.
+
+## Key Features
+
+The data on consumer debt information on various metrics of household indebtedness, including aggregate measures of credit market debt to disposable income, consumer credit, mortgage liabilities, non-mortgage loans, and other relevant indicators.
+The data cover a substantial time span, allowing for a longitudinal analysis of trends and patterns in consumer borrowing behavior over time.
+The interest rates dataset comprises time-series data on the Bank of Canada's key policy rate, commonly referred to as the overnight rate or the target for the overnight rate.
+Additionally, the dataset may include data on other short-term interest rates, such as the prime lending rate set by commercial banks, providing a comprehensive view of interest rate movements in the Canadian financial system.
+A preview of the merged datasets can be seen in @tbl-data.
+
+```{r}
+#| label: tbl-data
+#| tbl-cap: "A summary table of cleaned data"
+#| echo: false
+#| eval: true
+#| warning: false
+#| message: false
+
+data_overview <- head(analysis_data)
+
+# Use kable to create a nice-looking table
+kable(data_overview, caption = "Overview of Dataset")
+
+```
 
@@
 
-The plot visualizes the trends in the Credit Market Debt to Disposable Income ratio over time, spanning from 1990 to the present day.
+@fig-debt visualizes the trends in the Credit Market Debt to Disposable Income ratio over time, spanning from 1990 to the present day.
 The x-axis represents the date, while the y-axis depicts the Debt to Disposable Income Ratio, a key metric used to assess the financial health of households.
@@
 
-The plot illustrates the trends in short-term interest rates over time, spanning from 1990 to the present day.
+@fig-interest illustrates the trends in short-term interest rates over time, spanning from 1990 to the present day.
 The x-axis represents the date, while the y-axis depicts the level of short-term interest rates.
@@
 
-The graph @fig-regression visualizes the correlation between consumer debt and short-term interest rates.
+@fig-regression visualizes the correlation between consumer debt and short-term interest rates.
 

### 2024-04-20 paper/paper.qmd
Resolved merge conflicts

@@
 
-The model's primary explanatory variable is the short-term interest rate, represented by the IRSTCB01CAM156N variable, sourced from the dataset on short-term interest rates.
+The model's primary explanatory variable is the short-term interest rate, represented by the interest_rates variable, sourced from the dataset on short-term interest rates.
 This variable serves as a proxy for changes in the Bank of Canada's monetary policy stance, capturing variations in borrowing costs and the availability of credit in the economy.

### 2024-04-20 scripts/03-test_data.R

@@ -29,4 +29,3 @@ print(paste("Range of 'interest_rates':", interest_rates_range[1], "to", interes
 consumer_debt_range <- range(analysis_data$consumer_debt)
 print(paste("Range of 'consumer_debt':", consumer_debt_range[1], "to", consumer_debt_range[2]))
 
-

### 2024-04-20 other/datasheet/datasheet.qmd
Added the datasheet.

---
title: "Datasheet"
author: 
  - Abdul Ahad Qureshi
thanks: "Code and data are available at: https://github.com/abdulahadq29/debt."
date: today
date-format: long
format: pdf
number-sections: true
---

**Motivation**

1. *For what purpose was the dataset created? Was there a specific task in mind? Was there a specific gap that needed to be filled? Please provide a description.*
   - The dataset was created to analyze how changes in the Bank of Canada's interest rate policy affect consumer debt levels in Canada, aiming to understand the implications of monetary policy decisions on household indebtedness.
2. *Who created the dataset (for example, which team, research group) and on behalf of which entity (for example, company, institution, organization)?*
   - The dataset is compiled from Statistics Canada and Federal Reserve Economic Data (FRED), by researchers such as Abdul Ahad Qureshi for academic and policy analysis purposes.
3. *Who funded the creation of the dataset? If there is an associated grant, please provide the name of the grantor and the grant name and number.*
   - There is no specific information provided on funding or grants for the creation of the datasets from Statistics Canada and FRED; these datasets are maintained by governmental and institutional budgets.
4. *Any other comments?*
  - The study utilizes established, publicly available economic data to explore significant economic questions, aiding in policy formulation and economic understanding.

**Composition**

1. *What do the instances that comprise the dataset represent (for example, documents, photos, people, countries)? Are there multiple types of instances (for example, movies, users, and ratings; people and interactions between them; nodes and edges)? Please provide a description.*
	- The instances in the dataset represent economic and financial metrics (such as consumer debt levels and interest rates), not individuals or interactive entities.
2. *How many instances are there in total (of each type, if appropriate)?*
	- The exact number of instances (data points) is not specified but would typically include time-series data points spanning several years for each variable.
3. *Does the dataset contain all possible instances or is it a sample (not necessarily random) of instances from a larger set? If the dataset is a sample, then what is the larger set? Is the sample representative of the larger set (for example, geographic coverage)? If so, please describe how this representativeness was validated/verified. If it is not representative of the larger set, please describe why not (for example, to cover a more diverse range of instances, because instances were withheld or unavailable).*
	- The dataset comprises comprehensive data from national sources (Statistics Canada and FRED), representing macroeconomic trends rather than a sample from a larger dataset.
4. *What data does each instance consist of? "Raw" data (for example, unprocessed text or images) or features? In either case, please provide a description.*
	- Each instance consists of processed numerical data representing economic indicators like debt-to-income ratios and interest rates.
5. *Is there a label or target associated with each instance? If so, please provide a description.*
	- The primary target variable in the dataset is the Credit Market Debt to Disposable Income ratio, analyzed against interest rates.
6. *Is any information missing from individual instances? If so, please provide a description, explaining why this information is missing (for example, because it was unavailable). This does not include intentionally removed information, but might include, for example, redacted text.*
	- There is no specific mention of missing information; economic datasets typically undergo processing to handle any gaps or inconsistencies.
7. *Are relationships between individual instances made explicit (for example, users' movie ratings, social network links)? If so, please describe how these relationships are made explicit.*
	- Relationships are not between individual entries but are statistical correlations between different economic indicators (debt levels and interest rates) analyzed over time.
8. *Are there recommended data splits (for example, training, development/validation, testing)? If so, please provide a description of these splits, explaining the rationale behind them.*
	- Not applicable for this econometric study, as it uses complete datasets for comprehensive analysis without splitting into training/validation/testing sets.
9. *Are there any errors, sources of noise, or redundancies in the dataset? If so, please provide a description.*
	- No specific errors, noise, or redundancies are detailed, but economic datasets typically undergo cleaning to minimize these issues.
10. *Is the dataset self-contained, or does it link to or otherwise rely on external resources (for example, websites, tweets, other datasets)? If it links to or relies on external resources, a) are there guarantees that they will exist, and remain constant, over time; b) are there official archival versions of the complete dataset (that is, including the external resources as they existed at the time the dataset was created); c) are there any restrictions (for example, licenses, fees) associated with any of the external resources that might apply to a dataset consumer? Please provide descriptions of all external resources and any restrictions associated with them, as well as links or other access points, as appropriate.*
	- Relies on external data from Statistics Canada and FRED; these sources are stable and maintained but no absolute guarantee of constancy.
11. *Does the dataset contain data that might be considered confidential (for example, data that is protected by legal privilege or by doctor-patient confidentiality, data that includes the content of individuals' non-public communications)? If so, please provide a description.*
	- The dataset does not contain confidential information as it is based on aggregated economic data.
12. *Does the dataset contain data that, if viewed directly, might be offensive, insulting, threatening, or might otherwise cause anxiety? If so, please describe why.*
	- The data is neutral, consisting of economic indicators that are not offensive or personally threatening.
13. *Does the dataset identify any sub-populations (for example, by age, gender)? If so, please describe how these subpopulations are identified and provide a description of their respective distributions within the dataset.*
	- The dataset does not focus on sub-populations by demographics like age or gender, it's aggregated on a national level.
14. *Is it possible to identify individuals (that is, one or more natural persons), either directly or indirectly (that is, in combination with other data) from the dataset? If so, please describe how.*
	- It is not possible to identify individuals directly or indirectly as the data is aggregated and anonymized.
15. *Does the dataset contain data that might be considered sensitive in any way (for example, data that reveals race or ethnic origins, sexual orientations, religious beliefs, political opinions or union memberships, or locations; financial or health data; biometric or genetic data; forms of government identification, such as social security numbers; criminal history)? If so, please provide a description.*
	- The dataset does not contain sensitive data as it exclusively deals with macroeconomic indicators and financial rates.
16. *Any other comments?*
	- The dataset is essential for understanding economic impacts and policy implications but does not delve into individual-level data or sensitive personal details.

**Collection process**

1. *How was the data associated with each instance acquired? Was the data directly observable (for example, raw text, movie ratings), reported by subjects (for example, survey responses), or indirectly inferred/derived from other data (for example, part-of-speech tags, model-based guesses for age or language)? If the data was reported by subjects or indirectly inferred/derived from other data, was the data validated/verified? If so, please describe how.*
	- Data were derived from governmental economic databases (Statistics Canada, FRED), focusing on observable economic indicators.
2. *What mechanisms or procedures were used to collect the data (for example, hardware apparatuses or sensors, manual human curation, software programs, software APIs)? How were these mechanisms or procedures validated?*
	- Collected through official national data recording systems validated by respective government standards and protocols.
3. *If the dataset is a sample from a larger set, what was the sampling strategy (for example, deterministic, probabilistic with specific sampling probabilities)?*
	- The dataset is not a sample; it comprehensively includes all available data from the designated time periods.
4. *Who was involved in the data collection process (for example, students, crowdworkers, contractors) and how were they compensated (for example, how much were crowdworkers paid)?*
	- Data were collected by employees of government agencies without external compensation issues.
5. *Over what timeframe was the data collected? Does this timeframe match the creation timeframe of the data associated with the instances (for example, recent crawl of old news articles)? If not, please describe the timeframe in which the data associated with the instances was created.*
	- The data spans multiple years, reflecting ongoing economic conditions and policy changes as they occurred.
6. *Were any ethical review processes conducted (for example, by an institutional review board)? If so, please provide a description of these review processes, including the outcomes, as well as a link or other access point to any supporting documentation.*
	- As governmental data, collection and dissemination are typically subject to governmental review and regulations ensuring ethical standards.
7. *Did you collect the data from the individuals in question directly, or obtain it via third parties or other sources (for example, websites)?*
	- Data was obtained directly from established governmental economic databases, not from individuals directly.
8. *Were the individuals in question notified about the data collection? If so, please describe (or show with screenshots or other information) how notice was provided, and provide a link or other access point to, or otherwise reproduce, the exact language of the notification itself.*
	- No individual notifications necessary as the data involves aggregated economic indicators without personal data.
9. *Did the individuals in question consent to the collection and use of their data? If so, please describe (or show with screenshots or other information) how consent was requested and provided, and provide a link or other access point to, or otherwise reproduce, the exact language to which the individuals consented.*
	- Consent not applicable; the data comprises macroeconomic data publicly released for analysis and policy-making.
10. *If consent was obtained, were the consenting individuals provided with a mechanism to revoke their consent in the future or for certain uses? If so, please provide a description, as well as a link or other access point to the mechanism (if appropriate).*
	- Not applicable as the data does not involve personal or sensitive information requiring consent.
11. *Has an analysis of the potential impact of the dataset and its use on data subjects (for example, a data protection impact analysis) been conducted? If so, please provide a description of this analysis, including the outcomes, as well as a link or other access point to any supporting documentation.*
	- Likely conducted internally by the governmental bodies to ensure compliance with data protection laws, though specifics are not detailed in the document.
12. *Any other comments?*
	- The dataset focuses on economic metrics for macroeconomic analysis and policy assessment, without involving personal data collection or ethical concerns typical of datasets involving individual participants.

**Preprocessing/cleaning/labeling**

1. *Was any preprocessing/cleaning/labeling of the data done (for example, discretization or bucketing, tokenization, part-of-speech tagging, SIFT feature extraction, removal of instances, processing of missing values)? If so, please provide a description. If not, you may skip the remaining questions in this section.*
	- Yes, the data underwent cleaning and processing such as removing outliers, handling missing values, and standardizing variable formats for robust econometric analysis.
2. *Was the "raw" data saved in addition to the preprocessed/cleaned/labeled data (for example, to support unanticipated future uses)? If so, please provide a link or other access point to the "raw" data.*
	- The raw data is typically maintained by the original sources, Statistics Canada and FRED, which may be accessible for verification and comparison.
3. *Is the software that was used to preprocess/clean/label the data available? If so, please provide a link or other access point.*
	- The preprocessing likely used statistical software like R; however, specific scripts or tools used were not detailed but could be similar to those publicly available for econometric analysis.
4. *Any other comments?*
	- Preprocessing ensures the integrity and usability of the data for accurate economic analysis and policy evaluation.

**Uses**

1. *Has the dataset been used for any tasks already? If so, please provide a description.*
	- The dataset has been used to analyze the impact of the Bank of Canada's interest rate policy on consumer debt levels in Canada, focusing on economic research and policy implications.
2. *Is there a repository that links to any or all papers or systems that use the dataset? If so, please provide a link or other access point.*
	- No specific repository linking to papers or systems using the dataset was mentioned, but relevant studies might be found through academic databases or the websites of Statistics Canada and FRED.
3. *What (other) tasks could the dataset be used for?*
	- Beyond economic policy analysis, the dataset could be used for financial market prediction, academic studies in macroeconomics, or sector-specific analyses impacting consumer finance.
4. *Is there anything about the composition of the dataset or the way it was collected and preprocessed/cleaned/labeled that might impact future uses? For example, is there anything that a dataset consumer might need to know to avoid uses that could result in unfair treatment of individuals or groups (for example, stereotyping, quality of service issues) or other risks or harms (for example, legal risks, financial harms)? If so, please provide a description. Is there anything a dataset consumer could do to mitigate these risks or harms?*
	- The dataset’s aggregation and anonymization ensure no risk of unfair treatment or stereotyping of individuals; future users should continue using the data for aggregated macroeconomic analysis to avoid misinterpretation.
5. *Are there tasks for which the dataset should not be used? If so, please provide a description.*
	- The dataset should not be used for individual financial advice or predictions about specific personal financial outcomes, as it is designed for aggregate economic analysis.
6. *Any other comments?*
	- The dataset is a robust tool for understanding broad economic trends and should be used within the scope of its economic and policy-oriented purposes.

**Distribution**

1. *Will the dataset be distributed to third parties outside of the entity (for example, company, institution, organization) on behalf of which the dataset was created? If so, please provide a description.*
	- The dataset, derived from Statistics Canada and FRED, is already publicly available and distributed by these entities to anyone interested.
2. *How will the dataset be distributed (for example, tarball on website, API, GitHub)? Does the dataset have a digital object identifier (DOI)?*
	- Distributed online through websites and data portals of Statistics Canada and FRED, without specific mention of a DOI for the combined dataset used in the study.
3. *When will the dataset be distributed?*
	- The datasets from Statistics Canada and FRED are continually updated and available for access at any time through their respective platforms.
4. *Will the dataset be distributed under a copyright or other intellectual property (IP) license, and/or under applicable terms of use (ToU)? If so, please describe this license and/ or ToU, and provide a link or other access point to, or otherwise reproduce, any relevant licensing terms or ToU, as well as any fees associated with these restrictions.*
	- Data from Statistics Canada and FRED is generally available under open licenses that allow for wide usage, though specific terms can be found on their respective websites.
5. *Have any third parties imposed IP-based or other restrictions on the data associated with the instances? If so, please describe these restrictions, and provide a link or other access point to, or otherwise reproduce, any relevant licensing terms, as well as any fees associated with these restrictions.*
	- No additional third-party IP restrictions are imposed on the datasets from Statistics Canada and FRED beyond standard copyright notices applicable to government data.
6. *Do any export controls or other regulatory restrictions apply to the dataset or to individual instances? If so, please describe these restrictions, and provide a link or other access point to, or otherwise reproduce, any supporting documentation.*
	- No export controls or other regulatory restrictions specifically apply to these datasets, as they contain economic data intended for public research and policy analysis.
7. *Any other comments?*
	- Users should ensure compliance with the terms of use provided by Statistics Canada and FRED when using the data for research or other purposes.

**Maintenance**

1. *Who will be supporting/hosting/maintaining the dataset?*
	- The datasets are maintained by their respective sources, Statistics Canada and FRED (Federal Reserve Economic Data).
2. *How can the owner/curator/manager of the dataset be contacted (for example, email address)?*
	- Contact can typically be made through the official websites of Statistics Canada and FRED, where they provide customer service or contact links.
3. *Is there an erratum? If so, please provide a link or other access point.*
	- There's no specific erratum link provided; issues are generally handled through updates or notices directly on the respective websites.
4. *Will the dataset be updated (for example, to correct labeling errors, add new instances, delete instances)? If so, please describe how often, by whom, and how updates will be communicated to dataset consumers (for example, mailing list, GitHub)?*
	- Both Statistics Canada and FRED regularly update their datasets; updates are communicated via their websites or through official publications.
5. *If the dataset relates to people, are there applicable limits on the retention of the data associated with the instances (for example, were the individuals in question told that their data would be retained for a fixed period of time and then deleted)? If so, please describe these limits and explain how they will be enforced.*
	- Not applicable, as the dataset consists of aggregated economic data, not individual-level data.
6. *Will older versions of the dataset continue to be supported/hosted/maintained? If so, please describe how. If not, please describe how its obsolescence will be communicated to dataset consumers.*
	- Older versions of economic data are often archived and accessible for historical research through Statistics Canada and FRED websites.
7. *If others want to extend/augment/build on/contribute to the dataset, is there a mechanism for them to do so? If so, please provide a description. Will these contributions be validated/verified? If so, please describe how. If not, why not? Is there a process for communicating/distributing these contributions to dataset consumers? If so, please provide a description.*
	- Extensions or contributions to the official datasets are typically not applicable as these are controlled by national and federal institutions.
8. *Any other comments?*
	- The datasets are crucial resources for economic analysis and are maintained with high standards to ensure reliability and accuracy for public and academic use.

### 2024-04-20 paper/paper.qmd
added data source reference

@@
 These datasets provide the empirical foundation for investigating the relationship between changes in the Bank of Canada's interest rate policy and consumer borrowing behavior.
-The primary source of data on consumer debt levels is sourced from Statistics Canada.
-Complementing the dataset on consumer debt levels is information on short-term interest rates, obtained from FRED.
+The primary source of data on consumer debt levels is sourced from Statistics Canada [@StatCan].
+Complementing the dataset on consumer debt levels is information on short-term interest rates, obtained from FRED [@fred].
 

### 2024-04-20 paper/paper.qmd
updated files

@@
 ---
-title: "Impact of Bank of Canada's Monetary Policy on Consumer Debt Levels"
+title: "Analyzing the Impact of Bank of Canada's Monetary Policy on Consumer Debt"
 author: 
@@
 
+\newpage
 # References

### 2024-04-20 scripts/01-download_data.R
Update 01-download_data.R

@@ -1,7 +1,7 @@
 #### Preamble ####
 # Purpose: Downloads and saves the data from FRED and Statistics Canada.
 # Author: Abdul Ahad Qureshi
-# Date: 19 April 2024
+# Date: 18 April 2024
 # Contact: ahad.qureshi@mail.utoronto.ca
 # License: MIT
 

### 2024-04-20 scripts/00-simulate_data.R
Update 00-simulate_data.R

@@ -10,17 +10,16 @@
 library(tidyverse)
 
 
-# Assuming 'cleaned_data' is your dataset after cleaning and it contains
 # 'consumer_debt' as consumer debt and 'interest_rates' as interest rates.
 
 set.seed(123) 
-n <- nrow(cleaned_data) 
+n <- nrow(analysis_data) 
 beta_0 <- 200 
 beta_1 <- -5 
 sigma <- 50 
 
 # Simulate interest rates based on the range observed in dataset
-sim_interest_rates <- runif(n, min(cleaned_data$interest_rates), max(cleaned_data$interest_rates))
+sim_interest_rates <- runif(n, min(analysis_data$interest_rates), max(analysis_data$interest_rates))
 
 # Simulate consumer debt based on a linear model with added random noise
 sim_consumer_debt <- beta_0 + beta_1 * sim_interest_rates + rnorm(n, mean = 0, sd = sigma)
